Predicting political party affiliation from text

### 2016-06-21 manuscript/poltext.tex
Kleine Änderungen im gesamten Text und einige Umstellungen in results section

\documentclass[11pt]{article}
\usepackage{poltext2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

@@

\author{Felix Biessmann\thanks{~\tt felix.biessmann@gmail.com}\\
    \And
 Pola Lehmann\thanks{ ~{\tt pola.lehmann@wzb.eu} }\\
  Wissenschaftszentrum Berlin\\
\And 
Daniel Kirsch
 \And
  Sebastian Schelter \thanks{~\tt sebastian.schelter@tu-berlin.de}\\ TU Berlin
}

@@

\section{Introduction}
\label{sec:intro}
%
Analysis and classifications of political text is and has been a very important tool to generate political science data (Benoit et al. 2015). Traditionally such classifications are done by experts, who read and label the text of interest.\footnote{See for example the Manifesto Project (Budge et al. 2001; Klingemann et al. 2006), the Comparative Agendas Project or Poltext (Pétry/Duval 2015).} This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze. The growing field of automated text analysis, that allows the analysis of much more text in less time, is therefore of great interest to political scientists. Additionally automated text analyses allow for a more objective and replicable analysis of political text then human coders could achieve (Benoit et al. 2009)

A major problem with automated text analyses is generalisation to other text domains than that on which the system was trained. Human coders have read enough texts of different domains to detect political bias in a variety of contexts and styles. Machine learning algorithms however are naturally prone to poor generalisation performance if the training data is biased towards one text domain. Unfortunately good unbiased training data is difficult to obtain. One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data as it is a large source of text that can be clearly associated with a party. Here we examine to what extent these models generalise their to other text domains, such as party manifestos and texts from social media. 
We discuss the effects of text length and domain shift of text data and investigate some potential reasons for the differences in classification performance.

@@

In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results.

@@

\paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label. The protocols of plenary debates are available through the website of the German Bundestag \cite{bundestag}; an open source API was used to query the data in a cleaned and structured format \cite{bundestag-github}. Uninterrupted parts were treated as separate speech.

@@

\paragraph{Facebook post data}
For each party we crawled their facebook pages \cite{gruene-fb, spd-fb, cducsu-fb, linke-fb} and extracted the post text, excluding all comments and other information. Like the manifesto data, also these texts were very short. As aggregation per topics was not possible for this data, we aggregated the texts by splitting all texts into parts of 1000 words. In order to balance the classes we drew 50 random parts for each party. This procedure reduces the effect of text length and biased sampling of different parties when testing the classifier.

@@

The following section gives an overview of the results for all political bias prediction tasks. 
Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.

\subsection{Predicting political party affiliation}
When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained. The evaluation results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th Bundestag are listed in \autoref{tab:results_in-domain}.
These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament.

@@

\paragraph{Validation on manifesto data}
For out-of domain data obtained from manifesto data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}. This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}. When testing the classifier on the facebook data (results not shown), this drop is similar.

\paragraph{Length of texts influences prediction accuracy}
A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}. In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into political topics. The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0. This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.

@@

\subsection{Sentiment correlates with political power} 
We quantified the correlation between government membership as well as number of seats in the parliament with text sentiment. Our results in  \autoref{tab:sentiments} show that government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.

@@

%
\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
We quantified which words were preferentially used by each party by measuring the correlation of single words with the party label. Some unspecific stopwords are excluded. 
\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em Staatssekretaerin, Kanzlerin, Bundestagsfraktion}) and some words related to cutting of expenses ({\em Kuerzungen}).
\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit, Stabilitaet}).

\section{Conclusions and Limitations}\label{sec:conclusion}
Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain. These results encourage to use such systems as assistive technology for instance for human annotators in an active learning setting.

We find a strong effect of text length and text domain on the generalization performance of the classifier, as also shown in other studies on similar data \cite{Yu2008, Hirst2014}. The first effect, longer texts are easier to classify, makes intuitive sense. Also humans are worse at judging the political bias of shorter texts out of context: Benoit et al. (2015), for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders. However short texts are a realistic challenge for automated political bias prediction systems: Often political texts from social media data and other web media are indeed very short. But political education would benefit from automatic analyses of these data streams; these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data. \\
The second effect, drop in generalization performance on out-of-domain data, can be alleviated by aggregating texts into longer segments. In the case of party affiliation prediction, the out-of-domain classification is on par or even better than the prediction accuracy on in-domain data. However in the binary classification setting (government membership), aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain classification. We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.

>>>>

\documentclass[11pt]{article}
%\usepackage{poltext2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

@@

\author{Felix Biessmann\thanks{~\tt felix.biessmann@gmail.com}\\
    \And
 Pola Lehmann\thanks{ ~{\tt pola.lehmann@wzb.eu} }\\
  Wissenschaftszentrum Berlin für Sozialforschung\\
\And 
Daniel Kirsch
 \And
  Sebastian Schelter \thanks{~\tt sebastian.schelter@tu-berlin.de}\\ TU Berlin
}
% Sieht irgendwie komisch aus, mit nur zwei angegebenen affiliations, und meine ist ja auch so lang. Sollen wir die nicht in die Fussnote zur email Adresse packen?

@@

\section{Introduction}
\label{sec:intro}
%
Analysis and classifications of political text is and has been a very important tool to generate political science data \cite{Benoit.Forthcoming}. Traditionally such classifications are done by experts, who read and label the text of interest.\footnote{See for example the Manifesto Project, the Comparative Agendas Project or Poltext.} This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze. The growing field of automated text analysis, that allows the analysis of much more text in less time, is therefore of great interest to political scientists. Additionally automated text analyses allow for a more objective and replicable analysis of political text then human coders could achieve \cite{Benoit.2}.

A major problem with automated text analyses is generalisation to other text domains than that on which the system was trained \cite{Slapin.2014}. Human coders have read enough texts of different domains to detect political bias in a variety of contexts and styles. 
% Den vorangegangenen Satz würde ich streichen, weil ich mit der Aussage nicht übereinstimme. Das ist wahrscheinlich wieder der Unterschied der Disziplinen, ihr müsst rechtfertigen, warum Computer es können, wenn Menschen es können, wir müssen in der Regel rechtgertigen warum Menschen es können, wenn Computer es besser machen. 
Machine learning algorithms however are naturally prone to poor generalisation performance if the training data is biased towards one text domain. Unfortunately good unbiased training data is difficult to obtain. One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data as it is a large source of text that can be clearly associated with a party. Here we examine to what extent models trained on this data can be generalised to other text domains, such as party manifestos and texts from social media. 
We discuss the effects of text length and domain shift of text data and investigate some potential reasons for the differences in classification performance.

@@

In the following, \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results.

@@

\paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label. The protocols of plenary debates are available through the website of the German Bundestag \cite{bundestag}; an open source API was used to query the data in a cleaned and structured format \cite{bundestag-github}. Each uninterrupted part was treated as a separate speech.

@@

\paragraph{Facebook post data}
For each party we crawled their facebook pages \cite{gruene-fb, spd-fb, cducsu-fb, linke-fb} and extracted the post text, excluding all comments and other information. Like the manifesto data, these texts are very short. As aggregation per topic was not possible for this data, we aggregated the texts by splitting all texts into parts of 1000 words.

@@

The following section gives an overview of the results for all political bias prediction tasks. 
Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.\footnote{We used the speeches from the 17th legislative period for the first task, as this legislature is already completed and offers more data. We had to use the speeches from the 18th legislative period for the second task, because we only have recent facebook posts.}
% Frage: Wollen wir nicht schreiben, dass wir das erste auf 17 und 18 gerechnet haben, aber nur 17. reporten? Stimmt ja und macht die Ergebnisse vielleicht vertrauensvoller. Einen entsprechenden Satz hatte ich ja schonmal formuliert.

\subsection{Predicting political party affiliation}
\paragraph{... on in-domain-data}
When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained. The evaluation results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th Bundestag are listed in \autoref{tab:results_in-domain}.
These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament.

@@

\paragraph{... on out-of-domain-data}
% Validation on manifesto data hat mich hier irritiert, weil es in dem Absatz ja auch um die facebook Daten geht. 
For out-of domain data obtained from manifesto data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}. This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}. When testing the classifier on the facebook data (results not shown), this drop is similar.

\subsection{Length of texts influences prediction accuracy}
A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}. In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into political topics. The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0. This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.

@@

\subsection{The discriminatory effect between the parties}
\label{sec:discrim_effect}
An important question is, of course, whether the difference between  the parties relies on different policies or if the differentiating effect between the parties is based on other aspects of the text. To get more insight into what aspects are responsible for discrimanting between the parties, we analysed, first, whether the words used most commonly by each party are related to policies in line with the parties' ideologies and second, we examined whether sentiment has a discriminatory effect between government and opposition parties, too.

\paragraph{Correlations between words and parties}
We quantified which words were preferentially used by each party by measuring the correlation of single words with the party label. Some unspecific stopwords are excluded. We find clear differences between the parties and these difference are in line with the parties ideologies. 
\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em staatssekretaerin, kanzlerin, bundestagsfraktion}) and some words related to cutting of expenses ({\em kuerzungen}).
\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em wettbewerbsfaehigkeit, entwicklung}) and words related to security ({\em sicherheit, stabilitaet}).

\paragraph{Sentiment correlates with political power} 
We quantified the correlation between government membership as well as number of seats in the parliament with text sentiment. Our results in  \autoref{tab:sentiments} show that government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.

@@

%

\section{Conclusions and Limitations}\label{sec:conclusion}
Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain. These are encouraging results with respect to the use of such systems as assistive technology, for instance for human annotators in an active learning setting.

In line with other findings \cite{Yu2008, Hirst2014}, we find a strong effect of text length and text domain on the generalization performance of the classifier. The first effect, longer texts are easier to classify, makes intuitive sense. Also humans are worse at judging the political bias of shorter texts out of context: \cite{Benoit.Forthcoming}, for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders. However short texts are a realistic challenge for automated political bias prediction systems: Often political texts from social media data and other web media are indeed very short. But political scientists can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data. \\
The second effect, i.e. the drop in generalization performance on out-of-domain data, can be alleviated by aggregating texts into longer segments. In the case of party affiliation prediction, the out-of-domain classification is on a par or even better than the prediction accuracy on in-domain data. However in the binary classification setting (government membership), aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain classification. We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.
